The Anatomy of a Temporal Summarization System

### 2015-02-26 experiments.tex
latest revision changes

We train a salience model for each event using 1000 sentences randomly sampled
from the event's document stream. Kernel length scale parameters  for each
feature group are fit to maximize
the model's log likelihood using ???.
At test time, we predict a sentence's salience using the average predictions
of the 23 other models---essentially 24 fold cross validation approach.

@@

\subsection{Nugget Evaluation}
We also modify the evaluation used in the TREC TS track, 
where systems were evaluated on the containment of event nuggets. 
We match each system update to 0 or more nuggets, where an update matches
a nugget if their semantic similarity is above a threshold.  
Summaries are then evaluated on the precision and recall of the matches.
More formally,
\[ \mathrm{Prec.} = \frac{|S_n|}{|S|} \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}\] 
where S is the set of system updates, $S_n$ is the set of nuggets mapped updates in $S$,
$N$ is the set of event nuggets, and $|\cdot|$ is the size of the set.
A precision of 1 indicates that every sentence contains one unique nugget.
Because an update can map to more than one nugget, it is possible to receive a precision greater than 1. 
A recall of 1 indicates that the summary has cover all nugget information for the event; the maximum
attainable recall is 1.
Determining 
an optimal threshold is difficult so we evaluate at threshold values ranging
from .5 to 1, where values closer to 1 are more conservative estimates of
performance.

>>>>

We train a salience model for each event using 1000 sentences randomly sampled
from the event's document stream. Kernel length scale parameters  for each
feature group are fit to maximize
the model's log likelihood using ???.
At test time, we predict a sentence's salience using the average predictions
of the 23 other models---essentially a leave-one-out approach.

@@

\subsection{Average Gain and Comprehensiveness}

We treat the event's nuggets as unique units of information.
When a system adds an update to its summary, it is potentially adding some
of this nugget information. It would be instructive to know how much unique
and novel information each update is adding on average to the summary.
To that end, we define

\[ \mathrm{Avg. Gain} = \frac{|S_n|}{|S|}% \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}
\] 
where $S$ is the set of system updates, 
$S_n$ is the set of nuggets contained in $S$, and $|\cdot|$ is the number of
elements in the set.
To compute the set $S_n$ we match each system update to 0 or more nuggets, 
where an update matches a nugget if their semantic similarity is above 
a threshold. $S_n$ results from the unique set of nuggets matched.
Because an update can map to more than one nugget, it is possible to receive an
average gain greater than 1. An average gain of 1 would indicate that every
sentence was both relevant and contained a unique piece of information.

Additionally, we can use the nuggets to measure the completeness of an update 
summary. We define

\[ \mathrm{Comprehensiveness} = \frac{|S_n|}{|N|}\]

where $N$ is the set of event nuggets.
A comprehensiveness of 1 indicates that the summary has cover 
all nugget information for the event; the maximum
attainable comprehensiveness is 1.

Determining an optimal threshold to count matches is difficult so 
we evaluate at 
threshold values ranging
from .5 to 1, where values closer to 1 are more conservative estimates of
performance.
A qualitative assessment of matches 
suggests that semantic similarity values around .7 produce reasonable matches.
For example, the nugget 
\begin{blockquote}
watches and warnings in effect for much of the Mid-Atlantic states and 
southern New England 
\end{blockquote}
\noindent
matches with the update 
\begin{blockquote}
A tropical storm warning remains in effect for much of the coasts 
of North and South Carolina\end{blockquote}
\noindent
with a similarity of .7. 
The average semantic similarity of manual matches performed 
by TREC assessors was considerably lower at approximately .25, increasing
our confidence in the automatic matches in the .5-1 range.

An assessment of manual matchings done by TREC 
assessors revealed that the average cosine similarity

### 2015-02-26 introduction.tex
latest revision changes

\label{sec:introduction}
During crises, information is critical for first responders,
 crisis management organizations, and those caught in the event.  
When the event is significant, as in the case of Hurricane
Sandy, the amount of content produced by traditional news outlets,
government agencies, relief organizations, and social media can vastly
overwhelm those trying to monitor the situation. 
Crisis informatics\cite{?} is dedicated to finding methods for sharing the
right information in a timely fashion with relief organizations during a major crisis. 
Research in this field has focused on
human-in-the-loop approaches ranging from on the ground information gathering to
crowdsourced reporting and disaster management.

Multi-document summarization offers potential for enabling automatic updates of
relevant, salient information at regular intervals. It would provide
information even when human volunteers are unable to and would filter out
unnecessary and irrelevant detail. 
While methods for identifying,
tracking, and summarizing events from text based input have been explored
extensively
%KM - I think we should consider adding other event summarizers. 
(e.g.,
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}), 
these experiments were not developed to handle streaming data from the large
and heterogeneous environment of the modern web. 
These methods also rely heavily on redundancy which is suboptimal for time
sensitive domains where there is a high cost in delaying information.

%\cite{?} to enable people on the ground to update
%crisis interfaces with information about needs. Others have used crowd
%sourcing \cite{?},

%At times of crisis, however, social media can overwhelm current information
%systems with the quantity of information, much of which is irrelevant,
%unnecessary detail, or out of date.

%Neither were they developed to address the
%information needs that arise during  crisis
%situations. 
%KM - this last sentence sounds somewhat negative.
%there is still a need for robust and scalable 
%methods for automatic summarization.

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. 
Our system is better able to adapt to
dynamic changes in input volume which adversely 
effect methods of summarization that use either pairwise
comparisons as a proxy for salience or predict salience but
ignore redundancy in the input alone.

%We train a regression
%model to predict sentence salience and use these predictions to bias the
%formation of sentence clusters around more salient regions in the input space
%using affinity propagation (AP) clustering.  We adapt AP to incorporate
%the salience predictions
%as well as pairwise similarities among input sentences to identify
%\emph{exemplar} sentences, which we use as our summary output.

%ignoring features of importance that are intrinsic to the
%sentences themselves.

%KM - Chris - add a short overview on the kinds of things we measure and
%results so people are encouraged to read ahead. Keep it general.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about disaster to determine
salience. Thus, salience does not just represent importance within a set of
documents; it also represents both how typical this sentence is of the input 
event
type (i.e., disaster, hurricane, tornado) and whether it specifies information
about this particular disaster. 
Our feature representation includes a set of language models, one for each
disaster type, to measure typicality of the sentence for the current event 
type, the distance of mentioned locations from the center of
the disaster, and the change in word frequencies over the time of the event.
% to represent its likelihood of referring to the input disaster. 
%We also utilize the change in word frequencies as we process the stream 
%to measure the changing importance of updates over time.

Our approach achieves suprerior ROUGE scores compared to multiple baselines.
Additionally, we introduce a novel method for determining the 
coverage of atomic facts in a summary, and show our system's
superiority on this metric.

>>>>

\label{sec:introduction}
During crises, information is critical for first responders,
 crisis management organizations, and those caught in the event.  
When the event is significant, as in the case of Hurricane
Sandy, the amount of content produced by traditional news outlets,
government agencies, relief organizations, and social media can vastly
overwhelm those trying to monitor the situation. 
Crisis informatics \cite{?} is dedicated to finding methods for sharing the
right information in a timely fashion during such an event.
Research in this field has focused on
human-in-the-loop approaches ranging from on the ground information 
gathering to crowdsourced reporting and disaster management.

Multi-document summarization has the potential to assist the crisis 
informatics community. 
Automatic summarization could deliver
relevant and salient information at regular intervals, 
even when human volunteers are unable to. 
Perhaps more importantly it could help filter out
unnecessary and irrelevant detail when the volume of incoming information
is large. 
While methods for identifying,
tracking, and summarizing events from text based input have been explored
extensively
%KM - I think we should consider adding other event summarizers. 
(e.g.,
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}), 
these experiments were not developed to handle streaming data from a
 heterogeneous environment at web scale. 
These methods also rely heavily on redundancy which is suboptimal for time
sensitive domains where there is a high cost in delaying information.

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster, and integrates these predictions into
a clustering based multi-document summarization system. 
Not only does incorporating salience produce more relevant summaries,
our system is better able to adapt to
dynamic changes in input volume which adversely 
effect methods that use redundancy as a proxy for salience. 
%or predict salience but
%ignore redundancy in the input alone.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about the event to determine
salience. Thus, salience does not just represent importance within a set of
documents; it also represents both how typical a sentence is of the  
event
type (i.e., industrial accident, hurricane, riot) and whether it specifies 
information
about this particular event. 
Our feature representation includes a set of language models, one for each
event type, to measure the typicality of the sentence with regard to the 
current event, the distance of mentioned locations from the center of
the event, and the change in word frequencies over the time of the event.
While we ground these features in the domain of disaster, their templates
are generally applicable to many update summarization domains.
% to represent its likelihood of referring to the input disaster. 
%We also utilize the change in word frequencies as we process the stream 
%to measure the changing importance of updates over time.

Our approach achieves a statistically significant improvement in 
 ROUGE scores compared to multiple baselines.
Additionally, we introduce novel methods for estimating the average information
gain each update provides and how completely the update summary covers 
the event it is tracking; our system's updates contain more relevant
information on average than the competeing baselines.

### 2015-02-26 methods.tex
latest revision changes

In order to evaluate an update summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization (TS) Track.  We provide a brief overview of the problem.  Details on the formulation can be found in the track overview paper \cite{aslam2013trec}.

An update summarization system takes as input 
\begin{enumerate*}[label=\itshape\alph*\upshape)]
  \item a short query $\query$ defining the event to be tracked (e.g. `Hurricane Sandy'), 
  \item an event category $\eventcategory$ defining the type of event to be tracked (e.g. `hurricane'), 
  \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$,
  presented in temporal order, and \item an evaluation time period of interest,
      $(\stime,\etime)$.  \end{enumerate*} While processing documents
      throughout the time period of interest, the system must output sentences,
      known as updates, which are \emph{relevant} to the query,
      \emph{comprehensive} with respect to the event, \emph{novel} to the user,
      and \emph{timely} with respect to when the update occurred.  Precisely
      how to measure these properties will be discussed in Section
      \ref{sec:methods}.  In our version of this problem, we assume that that
      the system receives one batch of new sentence-segmented documents every
      hour throughout the period of interest.

We address disaster types such as terrorism and mass shootings (e.g., the 2012
shooting in Aurora, Col.), natural disasters (e.g., Hurricane Sandy), accidents
(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
Arab spring). The size and scope of the events varies considerably. Hurricane
Sandy, for example, effected multiple countries over weeks while the
2012 Aurora shooting was contained to a single location spanning several hours.

@@

Figure \ref{alg:temporal-summarization} outlines our general update
summarization algorithm.  At each hour, the system processes each input 
sentence batch $S_t$.
We predict the salience $P$ for all input sentences $s\in S_t$.
Next, we cluster $S_t$ using the AP clustering algorithm, biased by $P$,
obtaining a set of exemeplar sentences $E$. Finally we select a subset of $E$
to be updates and add those to our set of summary updates $U$.

\begin{algorithm}%[H]
 \KwData{ 
     $S_{\stime}, S_{\stime+1},\ldots,S_{\etime} $ --- time ordered sentence
     batches\\
 \KwResult{U --- a list of updates, i.e. the update summary} 
}
 ~\\
 Initialize empty list $U$ of updates\\
    \For{$t \gets \stime,\ldots,\etime $}{
        $P \gets \operatorname{PredictSalience}(S_t)$\\
        $E \gets \operatorname{APCluster}(P, S_t)$ \\
        $U_t \gets \operatorname{SelectUpdates}(E,P,U)$\\
        $U \gets U \cup U_t$
}
 \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
\end{algorithm}

\subsection{Salience Prediction}

\subsubsection{Model}
%KM - I think one of the words ``sentence'' below should be something else, btu
%not sure what you meant. Ah.. I think ``salience''?
%We adopt Gaussian process regression in order to predict sentence sentence
We adopt Gaussian process regression in order to predict sentence salience
\cite{rasmussen:gaussian-process-book}.  Gaussian process regressors are a
class of data-driven, non-parametric model generalizing the multi-variate
Gaussian to the infinite dimensional setting.  Gaussian processes are general
and are state of the art for many regression tasks.  A full treatment of
Gaussian process regression is beyond the scope of this article and can be
found in standard textbooks.

As a non-parametric, the Gaussian process relies on a  covariance matrix $\kernelMatrix$, measuring the affinity between pairs of instances, in our case candidate sentences.  In our experiments, we used a radial basis function (RBF) kernel.  Given two featurized sentences, $\features$ and $\features^\prime$, the RBF kernel is defined as,
\begin{align*}
        \kernelMatrixij{\features}{\features^\prime}&= \sigma^2 \exp\left(- \frac{1}{2} 
\sum_{i=1}^{\numfeatures} \frac{ (\featuresi{i}-\featuresi{i}^\prime)^2}{\ell_i^2} \right)
\end{align*}
where $\sigma$ and the $\ell_i$ are parameters we fit to our observed training data. The $\ell_i$ are feature dependent scaling parameters; once learned, they not only improve the accuracy of the model, but give us some introspection  into which features are more important.

\subsubsection{Training Data}
\ckcomment{Explain how we train the model here...}
%In order to train the salience model, we need to quantify the notion of salience. We assume for each event we have a set of atomic facts, or nuggets.

%
\begin{figure}[t!]
\begin{tabular}{| l |} 
\hline
\textbf{Basic Features}\\
$\cdot$ Sent. position (normalized by doc. length) \\
$\cdot$ Sent. length \\
$\cdot$ Ratio of punc. to non-punc. chars. \\
$\cdot$ Ratio of caps. to non-caps. chars. \\
$\cdot$ Ratio of lowercase to other chars. \\
$\cdot$ Ratio of uppercase to other chars. \\
$\cdot$ \# of caps. words (normalized by \# of words)\\
$\cdot$ \# of Person, Location, Org. Date, Number,\\
$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. \\
$\;\;$ tags (normalized by \# of words)\\
\hline
\textbf{Query Features}\\
$\cdot$ \% of query words covered by sent.\\
$\cdot$ Total query matches.\\
$\cdot$ Total event-type synonyms/hypernyms/ \\
$\;\;$ hyponyms coverage.\\
$\cdot$ Total event-type synonyms/hypernyms/ \\
$\;\;$ hyponyms matches.\\
\hline
\textbf{Language Model Features}\\
$\cdot$ Avg. token log probability (domain lang. \\
$\;\;$ model)\\
$\cdot$ Avg. token log probability (background \\
$\;\;$ lang. model)\\
\hline
\textbf{Geo-tag Features}\\
$\cdot$ Median document distance to nearest  \\
$\;\;$ location cluster (current hour).\\
$\cdot$ Median document distance to nearest  \\
$\;\;$ location clusters (previous hour).\\
$\cdot$ Distance of first location in doc. to nearest \\
$\;\;$ location cluster (current hour).\\
$\cdot$ Distance of first location in doc. to nearest \\
$\;\;$ location cluster (previous hour).\\
\hline
\textbf{Temporal Features}\\
$\cdot$ Avg. tf-idf at current time.\\
$\cdot$ Change in avg. tf-idf since previous hour \\
$\;\;$  (up to 24 hours).\\
$\cdot$ Time since query/event start.\\
\hline
\end{tabular}
\caption{Salience model features.}
\end{figure}

\subsubsection{Features}
We want our model to be predictive across different kinds of events so we avoid lexical features.  Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.  These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.

@@

We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.  
Since training is one on grammatical English, some of these features help
to downweight sentences that are ungrammatical (e.g., have too many capitalized words or are too short).
Others help to more heavily weight important sentences (e.g., that appear in
prominent positions such as paragraph initial or article initial).

@@

\subsection{Affinity Propagation Clustering}
\ckcomment{Explain AP and how salience gets plugged in here}
\subsection{Update Selection}
\ckcomment{Explain thresholds}

>>>>

%In order to evaluate an update summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization (TS) Track.  We provide a brief overview of the problem.  Details on the formulation can be found in the track overview paper \cite{aslam2013trec}.

An update summarization system takes as input 
\begin{enumerate*}[label=\itshape\alph*\upshape)]
  \item a short query defining the event to be tracked (e.g. `Hurricane Sandy'), 
  \item an event category defining the type of event to be tracked (e.g. `hurricane'), 
  \item a stream of time-stamped documents %, $(\doc_0, \doc_1,\ldots,\doc_T)$,
  presented in temporal order, and \item an evaluation time period of interest.
    %  $(\stime,\etime)$.  
      \end{enumerate*} While processing documents
      throughout the time period of interest, the system must output sentences,
      known as updates.
      Updates should not only be relevant to the event but also salient,
      i.e. worthy of
      reporting.
      For most domains, updates also need to be 
      timely and novel, i.e. users need to be informed of changes in the 
      event as quickly as possible.
      Finally, in aggregate the updates must comprehensively summarize the 
      event.

In order to quantify the salience of a sentence with respect to an event,
we assume the event can be represented as a set of atomic facts, or nuggets.
Nuggets are brief and important text snippets that should be conveyed
by an ideal update summary (see Figure~\ref{fig:nuggets} for examples).
\setlength{\fboxsep}{10pt}
\begin{figure}
    \fbox{\parbox{6.9cm}{
        $\cdot$ hurricane force wind warnings are in effect from Rhode Island Sound to Chincoteague Bay\\

$\cdot$ Obama declared an emergency in Maryland and signed an order authorized the Federal Emergency Management Agency to aid in disaster relief\\

        $\cdot$ over 5000 commercial airline flights scheduled for October 28 and October 29 were cancelled
 }}
 \caption{Example nuggets from Hurricane Sandy.\label{fig:nuggets}}
\end{figure}

Given an update $u$ and a set of nuggets $N$ we define the salience of 
$u$ as,
\[\operatorname{salience}(u) = \operatorname{max}_{n \in N} 
\operatorname{sim}(u, n) \]
where $\operatorname{sim}(\cdot)$ is the semantic similarity \cite{?} (i.e.
the cosine similarity of latent vectors associated with the update and 
nugget text). Since we will not have nuggets for new events, we train
a regression model to predict salience using features derived from 
the updates themselves
and the document stream.

     % \ref{sec:methods}.  In our version of this problem, we assume that that
     % the system receives one batch of new sentence-segmented documents every
     % hour throughout the period of interest.
\subsection{Update Summarization}

Our system archictecture follows a simple pipeline design where each
stage provides an additional level of processing or filtering of the input
sentences.
At the highest level, we begin with an empty update summary $U$.
At each hour we receive a new batch of sentences $b_t$ from the stream
and perform the following actions:
\begin{enumerate}
    \item predict the salience of sentences in $b_t$ (Section~\ref{sec:salpred}),
  \item select a set of exemplar sentences (by combining clustering with 
      salience predictions) (Section~\ref{sec:exsel}),
  \item add the most novel and salient exemplars to $U$ (Section~\ref{sec:upsel}).
\end{enumerate}

The resultant list of updates $U$ is our summary of the event.
Step 2 uses the affinity propagation algorithm to incorporate the salience
predictions from step 1. For this reason we dub our summarization 
system the \textbf{AP+Salience} model.

@@

%Figure \ref{alg:temporal-summarization} outlines our general update
%summarization algorithm.  At each hour, the system processes each input 
%sentence batch $S_t$.
%We predict the salience $P$ for all input sentences $s\in S_t$.
%Next, we cluster $S_t$ using the AP clustering algorithm, biased by $P$,
%obtaining a set of exemeplar sentences $E$. Finally we select a subset of $E$
%to be updates and add those to our set of summary updates $U$.
%
%\begin{algorithm}%[H]
% \KwData{ 
%     $S_{\stime}, S_{\stime+1},\ldots,S_{\etime} $ --- time ordered sentence
%     batches\\
% \KwResult{U --- a list of updates, i.e. the update summary} 
%}
% ~\\
% Initialize empty list $U$ of updates\\
%    \For{$t \gets \stime,\ldots,\etime $}{
%        $P \gets \operatorname{PredictSalience}(S_t)$\\
%        $E \gets \operatorname{APCluster}(P, S_t)$ \\
%        $U_t \gets \operatorname{SelectUpdates}(E,P,U)$\\
%        $U \gets U \cup U_t$
%}
% \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
%\end{algorithm}

\subsection{Salience Prediction}
\label{sec:salpred}

\subsubsection{Model}

Nuggets represent ground truth data which would not be available at test time;
we need to find some function $f : \mathbb{R}^d \rightarrow \mathbb{R}$ 
such that 
\[ \operatorname{salience}(s) = f(x_s) \]
where $x_s$ is a $d$-dimensional feature vector for sentence $s$. 
Given that we have many thousands of sentences of varying relevance to each
event in our stream and salience scores are derived in an unsupervised fashion
(assuming event nuggets are known), we can generate large amounts of training
data for our model by simply sampling sentences 
from the stream and computing the maximum
semantic similarity to the nugget set.

To learn $f$ and predict salience,
we adopt Gaussian process (GP) regression 
\cite{rasmussen:gaussian-process-book}.  GP regressors are a
class of data-driven, non-parametric model generalizing the multi-variate
Gaussian to the infinite dimensional setting.  GP's are general
and are state of the art for many regression tasks.  
%A full treatment of
%Gaussian process regression is beyond the scope of this article and can be
%found in standard textbooks.

GP's are non-parametric and rely on a covariance matrix $\kernelMatrix$, measuring the similarity between pairs of instances, in our 
case sentences.  In our experiments, we used a radial basis 
function (RBF) kernel.  Given two featurized sentences, 
$\features$ and $\features^\prime$, the RBF kernel is defined as,
\begin{align*}
        \kernelMatrixij{\features}{\features^\prime}&= \sigma^2 \exp\left(- \frac{1}{2} 
\sum_{i=1}^{\numfeatures} \frac{ (\featuresi{i}-\featuresi{i}^\prime)^2}{\ell_i^2} \right)
\end{align*}
where $\sigma$ and the $\ell_i$ are parameters we fit to our observed training data. The $\ell_i$ are feature dependent scaling parameters; once learned, they not only improve the accuracy of the model, but give us some introspection  into which features are more important.
Kernel matrices are also closed under addition, i.e. the sum of kernels
is also a kernel; we exploit this fact by organizing our features into 
meaningful groups with their own RBF kernel using the sum of their kernels 
to parameterize the model.

\begin{figure}[t!]
\begin{tabular}{| l |} 
\hline
\textbf{Basic Features}\\
$\cdot$ Sent. position (normalized by doc. length) \\
$\cdot$ Sent. length \\
$\cdot$ Ratio of punc. to non-punc. chars. \\
$\cdot$ Ratio of caps. to non-caps. chars. \\
$\cdot$ Ratio of lowercase to other chars. \\
$\cdot$ Ratio of uppercase to other chars. \\
$\cdot$ \# of caps. words (normalized by \# of words)\\
$\cdot$ \# of Person, Location, Org. Date, Number,\\
$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. \\
$\;\;$ tags (normalized by \# of words)\\
\hline
\textbf{Query Features}\\
$\cdot$ \% of query words covered by sent.\\
$\cdot$ Total query matches.\\
$\cdot$ Total event-type synonyms/hypernyms/ \\
$\;\;$ hyponyms coverage.\\
$\cdot$ Total event-type synonyms/hypernyms/ \\
$\;\;$ hyponyms matches.\\
\hline
\textbf{Language Model Features}\\
$\cdot$ Avg. token log probability (domain lang. \\
$\;\;$ model)\\
$\cdot$ Avg. token log probability (background \\
$\;\;$ lang. model)\\
\hline
\textbf{Geo-tag Features}\\
$\cdot$ Median document distance to nearest  \\
$\;\;$ location cluster (current hour).\\
$\cdot$ Median document distance to nearest  \\
$\;\;$ location clusters (previous hour).\\
$\cdot$ Distance of first location in doc. to nearest \\
$\;\;$ location cluster (current hour).\\
$\cdot$ Distance of first location in doc. to nearest \\
$\;\;$ location cluster (previous hour).\\
\hline
\textbf{Temporal Features}\\
$\cdot$ Avg. tf-idf at current time.\\
$\cdot$ Change in avg. tf-idf since previous hour \\
$\;\;$  (up to 24 hours).\\
$\cdot$ Time since query/event start.\\
\hline
\end{tabular}
\caption{Salience model features.}
\end{figure}

\subsubsection{Features}
We want our model to be predictive across different kinds of events so we avoid lexical features.  Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.  
%These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.

@@

We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.  
Since training is done on grammatical English, some of these features help
to downweight sentences that are ungrammatical (e.g., have too many capitalized words or are too short).
Others help to more heavily weight important sentences (e.g., that appear in
prominent positions such as paragraph initial or article initial).

@@

\subsection{Exemplar Selection}
\label{sec:exsel}

We combine the output of our salience prediction model with the affinity
propagation algorithm to identify a set of exemplar sentences 
for each input batch. 
Affinity propagation (AP) is a clustering algorithm
that identifies a subset of datapoints as exemplars and forms clusters
by assigning the remaining points to one of the exemplars. AP can be seen
as a generalization of other $k$-centers algorithms that attempts to 
maximize the net similarity objective 
\[ \mathcal{S} = \sum_{i : i \neq e_i}^n \operatorname{sim}(i,e_i) 
+ \sum_{i : i = e_i}^n \operatorname{pref}(e_i)  \]
where $e_i$ is the exemplar of the $i$-th data point, and functions
$\operatorname{sim}$ and $\operatorname{pref}$ express the pairwise similarity
of data points and the apriori preference of a data point to be an exemplar
respectively. There are very 
few restrictions on the nature of the similarities and preferences
other than that they be real-valued.

AP has several useful properties and interpretations. Chiefly, the number
of clusters $k$ is not a model hyper-parameter. Given that our task requires
clustering many batches of streaming data, searching for an optimal $k$ 
would be computationally prohibitive. With AP, $k$ is determined by the
similarities and preferences of the data. Generally lower preferences will
result in fewer clusters.

In our summarization system we use the output of our salience model as 
the preferences, i.e.
$\operatorname{salience}(s) = \operatorname{pref}(s)$; for the similarity 
function we use semantic similarity. Recall that $\operatorname{salience}(s)$
is a prediction of the semantic similarity of $s$ to information about the 
event be summarized, i.e. the set of event nuggets.
Intuitively, when maximizing objective function $\mathcal{S}$, AP must balance
between best representing the input data and representing the most salient
input. Additionally, when the level of input is high but the salience
predictions are low, the preference term will guide AP toward a solution
with fewer clusters; vice-versa when input is very salient on average but
the volume of input is low. The adaptive nature of our model differentiates
our method from most out update summarization systems.

\subsection{Update Selection}
\label{sec:upsel}

After exemplar sentences have been identified, two more filters are applied 
before updating the summary. The exemplars are examined
sequencially in order of highest salience. First, a salience threshold 
is applied, and any exemplars whose salience is less than
$\lambda_{sal}$ are discarded. Second, a similarity threshold 
is applied, where the exemplar is discarded if its
maximum semantic similarity to any previous updates in the summary is
greater than $\lambda_{sim}$. If the exemplar passes these two filters,
it is selected as an update and added to the summary.

### 2015-02-26 relatedwork.tex
latest revision changes

%\fdadd{
Several researchers have recognized the importance of summarization during
natural disasters.  Guo \textit{et al.} developed a system for detecting
novel, relevant, and comprehensive sentences immediately after a natural
disaster \cite{qi:temporal-summarization}.  The method uses a model of
sentence relevance and novelty in order to select appropriate updates.
Training data for regression targets is automatically generated from
retrospective Wikipedia data.  The system is evaluated on news documents
related to 197 natural and human disasters from 2009 to 2011 using variants of
Rouge modified to capture novelty, relevance, and comprehensiveness
\cite{lin2004rouge}.  Wang and Li present a clustering-based approach to
efficiency detect important updates during natural disasters
\cite{wang:update-summarization}.  The algorithm works by hierarchically
clustering sentences online, allowing the system to output a more expressive
narrative structure than Guo \textit{et al.}.  The method is evaluated on
official press releases related to Hurricane Wilma  in 2005 using Rouge score
between the system summary and a manually generated target summary.

\fdcomment{Glasgow temporal summarization system \cite{mccreadie:temporal-summarization}.}

>>>>

%\fdadd{
Several researchers have recognized the importance of summarization during
natural disasters.  Guo \textit{et al.} developed a system for detecting
novel, relevant, and comprehensive sentences immediately after a natural
disaster \cite{qi:temporal-summarization}.  The method uses a model of
sentence relevance and novelty in order to select appropriate updates.
Training data for regression targets is automatically generated from
retrospective Wikipedia data.  The system is evaluated on news documents
related to 197 natural and human disasters from 2009 to 2011 using variants of
Rouge modified to capture novelty, relevance, and comprehensiveness
\cite{lin2004rouge}.  Wang and Li present a clustering-based approach to
efficiency detect important updates during natural disasters
\cite{wang:update-summarization}.  The algorithm works by hierarchically
clustering sentences online, allowing the system to output a more expressive
narrative structure than Guo \textit{et al.}.  The method is evaluated on
official press releases related to Hurricane Wilma  in 2005 using Rouge score
between the system summary and a manually generated target summary. In addition
to natural disasters, we evaluate our system on a variety of event types 
 such as terrorism and mass shootings (e.g., the 2012
shooting in Aurora, Col.), accidents
(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
Arab spring). The size and scope of the events varies considerably. Hurricane
Sandy, for example, affected multiple countries over weeks while the
2012 Aurora shooting was contained to a single location spanning several hours.

%\fdcomment{Glasgow temporal summarization system \cite{mccreadie:temporal-summarization}.}

### 2015-02-26 results.tex
latest revision changes

@@

\begin{figure*}
  \includegraphics[width=\textwidth]{rouge-time.png}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
\end{figure*}

>>>>

Figure~\ref{fig:nuggets} shows the average gain across a range of 
similarity thresholds, where thresholds closer to 1 are more conservative
estimates. The ranking of the systems remains relatively constant across the 
sweep with 
AP+Salience and AP+Salience TR attaining comparable performance, as well
as beating all baseline systems.
Predicting salience in general is helpful for keeping a summary on topic as
the Rank by salience approach out performs the clustering only approaches
on average gain.

When looking at the comprehensiveness of the summaries AP outperforms
both AP+Salience variants. The compromise encoded in the AP+Salience objective
function, between being representitive and being salient, is seen clearly here
where the performance of the AP+Salience methods is lower bounded by 
the salience focused Rank
system and upper bounded by the clustering only AP system.

@@

\begin{figure}
  \includegraphics[width=3.2in,height=5in]{rouge-time.png}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
\end{figure}

\begin{figure}
  \includegraphics[width=3.2in,height=5in]{nugget-metrics.png}
\caption{Nugget performance.}
\label{fig:nperf}
\end{figure}
